%
%
\documentclass[11pt, letterpaper]{article}
\usepackage[dvips]{graphicx}
\usepackage{epsfig}
\usepackage{rotating}
\usepackage{times}
\pagestyle{empty}

%
% GET THE MARGINS RIGHT, THE UGLY WAY
%
\topmargin 0.0in
\textwidth 6.5in
\textheight 9.0in
\columnsep 0.25in
\oddsidemargin 0.0in
\evensidemargin 0.0in
\headsep 0.0in
\headheight 0.0in

\title{Current PVFS2 status}
\author{ PVFS2 Development Team }
\date{ Last Updated: September 2003 }

%
% BEGINNING OF DOCUMENT
%
\begin{document}
\maketitle

\tableofcontents

\newpage

\thispagestyle{empty}

\setlength{\parindent}{0.0cm}

\section{Introduction}

This document describes the current status of PVFS2 development.  This document
only includes issues related to functionality or correctness.  No 
performance optimizations are listed for now.

\section{Known limitiations and missing features}

This section lists file system limitations for which we have a known
solution or plan.

\begin{itemize}
\item efficient conversion of MPI datatypes to PVFS2 datatypes in ROMIO
\item hooks for tuning consistency semantics
\item hooks for controlling distribution and distribution parameters
\item standardizing error code format
\item integration of user level buffer cache
\item eliminating memory leaks
\item consistent error handling in client and server state machines
\item simple failover plan
\end{itemize}

\section{Experimental features}

These are features that are implemented but have not been thoroughly tested.

\begin{itemize}
\item GM network support
\item IB network support
\end{itemize}

\section{Placeholder / depricated code}

These parts of the code have a working implementation, but we intend to
replace them as time permits.

\begin{itemize}
\item ``contig'' request encoder implementation
\item pvfs2-client implementation
\end{itemize}

\section{Open issues}

The items on this list are known problems that have not been resolved.

\begin{itemize}
\item access control / security
\item how to manage client side configuration (fstab information)
\item support for 2.4 series kernels
\item how to add file systems to an existing system interface run time instance (proper vfs bootstrapping)
\item managing server configuration files
\item nonblocking I/O at system interface
\item how to handle I/O failures, unposting, etc.
\item TCP module scalability
\item extended attributes
\item redundancy
\end{itemize}

\section{Good examples}

This section points out specific areas of the code that demonstrate 
best practice for PVFS2 development.

\begin{itemize}
\item ?
\end{itemize}

\end{document}

